% !TEX encoding = UTF-8 Unicode

\section{\large Summary}

%%%%%%%%%%%%%%%%%%%%

\subsection{Conclusion}

\begin{frame}[fragile,t]{Conclusion}
	\begin{block}{For the System}
    \begin{itemize}
    \item The system is complete, encapsulated and user-friendly.
    \item The system performs well in adaptive predictions, 
    	all actual returns fall in the $95.45\%$ confidence interval.
    \item Time lag effects exist, so sample reweighting is necessary to improve the performance.
    \end{itemize}
    \end{block}
    
    \begin{block}{For Empirical Analyses}
    \begin{itemize}
    \item Data populations with longer observation periods tend to 
    	provide more accurate hidden states anlysis results then short-term data.
    \item Data populations with higher observation frequencies
		tend to outperform those with lower frequencies w.r.t.\,prediction.
    \end{itemize}
    \end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%

\subsection{Future Work}

\begin{frame}[fragile,t]{Future Work}
	\begin{block}{Time Lag Effects}
    \begin{itemize}
    \item \textbf{Rolling window method}, more popular in the industry,
    	the size of the window should be optimized.
    \item \textbf{Exponential weighted expectation maximization (EWEM) algorithm},
    	proposed in \cite{Zhang:2005tp}, more complex to implement.
    \end{itemize}
    \end{block}
    
    \begin{block}{Particle Filter}
    \begin{itemize}
    \item Standard tool for HMM estimation when the latent variable is continuous,
    	one of the most famous cases being Kalman filter.
    \item PF can incorporates economic meanings into values of the latent variable while currently 
    	we only consider the hidden states as categorical.
    \end{itemize}
    \end{block}
\end{frame}