%%==================================================
%% thanks.tex for SJTU Bachelor Thesis
%% version: 0.5.2
%% Encoding: UTF-8
%%==================================================

\begin{thanks}

I would like to express my great gratitude to my parents, teachers and friends.
Thank you for backing me up all the time.
It is never tough to come up with an undergraduation thesis,
but it is never easy to keep fighting until the last second of my college life even I am graduating.
Life has been especially hard in the last year,
and thanks to you all because I know I would be confused without your advice and encouragements.

Thanks to Prof. Jianzhong Lin for being the my supervisor of the thesis, 
and for spending time reading the thesis and providing suggestions.

Thanks to my parents for never pushing me even slightly.
Thank you for telling me to stop and take a break whenever I am working hard.
I know I would never stop, 
yet I still need this to find solace when I am too frustrated to rest.

Special thanks to Prof. Jinglai Li for introducing me to the realm of academic research,
and for all the supports during my graduate school applications.
Special thanks to Qun Zhao, 
a true friend and a great mentor,
for all kinds of advice and help on career path and academic research and 
other things that are too numerous to mention.
Special thanks to my competitors during all exams, interviews and contests.
Thank you for always reminding me that there is still a long long way to go,
to become a better person and to do more things.

Finally, special thanks to myself,
for never giving up and never having the thought to give up,
for all the fun and excitement during the life adventure,
and for all the passion that keeps me fighting and looking for new funs.

Life sucks when you are ordinary.
So be more than that,
beat whatever stands between yourself and your goals,
especially your past selves,
the vulnerable and weak ones that hinder you from going further.
If you should fail, you fail,
but never stop trying.
If you would live, you live,
and always live better, 
long and prosper.

% Special thanks to my future girlfriend, 
% if there should be one,
% for not showing up during the last year so that I could be fully concentrated on whatever I was doing.
% Now it is time to show yourself.
% 
% Special and great thanks to Fat Huang, Big Huang, Small Huang, 
% Small Hua, Small Hui, Small Bai, Small Hei, et al., 
% from PYG Library, without whom I would never have enjoyed the last year so much.
%
% Life sucks when you are ordinary.
% Fuck life, and enjoy it.

\end{thanks}

